---
title: "Experiment-5"
author: "siju.swamy@saintgits.org"
date: "14/07/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
> Study of confidence intervals How to compute confidence intervals for the mean when the standard deviation is known.

## Study of Confidence Intervals for means of Large and Small Samples

### Calculate Confidence Interval in R – Normal Distribution

confidence interval in r confidence interval r confidence intervals confidence interval in r linear regression confint r histogram with confidence interval in r confidence intervals in r plot confidence interval in r 
 confidence interval function percentile confidence interval in r confidence interval for variance in r how to calculate 95 percent confidence interval in r confidence confidence interval function in r
 
Given the parameters of the population proportion distribution and sample standard deviation, generate the bootstrap confidence interval. In this situation, we’re basically using r like an error interval calculator… Using the 95 percent confidence level and confidence coefficient function, we will now create the R code for a confidence interval. What does a 95 percent confidence interval mean? Essentially, a calculating a 95 percent confidence interval in R means that we are 95 percent sure that the true probability falls within the confidence interval range that we create in a standard normal distribution.

$$CI=\bar{X}\pm t_{0.975}SE$$

```{r}
# Calculate Confidence Interval in R for Normal Distribution
# Confidence Interval Statistics
# Assume mean of 12
# Standard deviation of 3
# Sample size of 30
# 95 percent confidence interval so right tail at 0.975

xbar <- 12
stddev <- 3
n <- 30
error <- qnorm(0.975)*stddev/sqrt(n)
lower_bound <- xbar - error
upper_bound <- xbar + error
lower_bound
upper_bound
```

### Calculate Confidence Interval in R – t Distribution

For experiments run with small sample sizes it is generally inappropriate to use the standard normal distribution or normal approximation. For more accurate small sample hypothesis testing a student T distribution is the correct choice for this environment. A t confidence interval is slightly different from a normal or percentile approximate confidence interval in R. When creating a approximate confidence interval using a t table or student t distribution, you help to eliminate some of the variability in your data by using a slightly different base dataset binomial distribution.

R can support this by substituting the qt function for the qnorm function, as demonstrated below…. assume we are working with a semi large sample size of 15. You will need to tell the qt function the degrees of freedom as a parameter (should be n-1).

```{r}
# Calculate Confidence Interval in R for t Distribution
# t test confidence interval
# Assume mean of 12
# Standard deviation of 3
# Sample size of 15
# 95% confidence interval so the right tail at .975

xbar <- 12
stddev <- 3
n <- 30
error <- qt(0.975,df=n-1)*stddev/sqrt(n)
lower_bound <- xbar - error
upper_bound <- xbar + error
lower_bound
upper_bound
```

